A Horrocks-type theorem for even orthogonal $K_2$

### 2020-07-23 horrocks.tex
Replace sharp by acute

The following proposition is the main result of this subsection and also a direct generalization of~\cite[Lemma~3.3]{Tu83}.
\begin{prop} \label{lemma33} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the homomorphism $i_n\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$, induced by the natural embedding of generators, is an isomorphism. Consequently, the obvious homomorphism $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{prop}
\begin{proof}
 We need to construct a homomorphism $j_n$ which would be the inverse of $i_n$. 
 We start with a homomorphism $\widetilde{j}_n^\Phi \colon F\langle \mathcal{X}^\Phi_{\leq n+1} \rangle \to \St^{\leq n}(\Phi, B)$ defined via
 \[ \widetilde{j}^{\Phi}_n(x_\alpha(at^k)) = \begin{cases} x_\alpha(at^k), & k\leq n; \\
      [x_{\alpha - \beta} (N_{\alpha-\beta, \beta} \cdot at^{k-1}), x_{\beta}(t)], & k = n+1, \end{cases} \]
 where $\beta$ is any root of $\Phi$ forming a sharp angle with $\alpha$.
 A standard argument, cf. \cite[Proposition~1.1]{Re75} or~\cite[Proposition~3.2.2]{RS76}, shows that $\widetilde{j}^\Phi_n$ does not depend on the choice of $\beta$.

@@

For the rest of this subsection $M$ is the maximal ideal $A$.
We denote by $\beta$ another fixed root of $\Phi$ forming a sharp angle with $\alpha$ (i.\,e. $\langle \alpha, \beta \rangle = 1$). We denote by $\Psi$ the subsystem of type $\rA_2$ generated by $\alpha$ and $\beta$.
\begin{df}\label{def:Kab} 
Fix a root $\delta \in \Phi \setminus \Psi$. Define $\widetilde{K}(\alpha, \beta)$ as the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following 5 families of elements parametrized by $\xi \in A[X]$, $f\in M[X]$:
 \begin{enumerate}[label=(K\arabic*)]
  \item $z_\gamma(Xf, \xi)$, for all $\gamma \in \Phi \setminus \Psi$;
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $x_{\alpha-\beta}(Xf)$, $x_{\beta-\alpha}(Xf)$;
  \item $c_{\delta}(f, X\xi)$. \end{enumerate} \end{df}
\begin{prop} \label{K-a-b} One has \[K(A[X], M[X]) = \widetilde{K}(\alpha, \beta) \cdot X_{-\alpha}(M \cdot X) \cdot X_{-\beta}(M \cdot X).\] 
\end{prop} 
\begin{proof}
It is clear that $K(A[X], M[X]) \supseteq \widetilde{K}(\alpha, \beta) \cdot X_{-\alpha}(M \cdot X) \cdot X_{-\beta}(M \cdot X)$. Let us prove the reverse inclusion. Fix an element $g\in K(A[X], M[X])$. We need to show that $g$ lies in the specified product of subgroups.

@@

It remains to verify the assertion for the generators $x_{-\alpha}(X^2f) $ of type K2.
By~\cref{K-a-b-cor} we may assume that the root $\delta$ in~\cref{def:Kab} forms a sharp angle with $\alpha$, i.\,e. $\alpha-\delta \in \Phi$. Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = \delta-\alpha$, $\beta = -\delta$ into~\cref{Zrels}\ref{Z5} we obtain that
\begin{multline*} z_{-\alpha}(X^2f, aX^{-1}) = x_{\delta-\alpha}(\epsilon Xf) \cdot x_{\delta}(-af) \cdot x_{-\delta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{\delta-\alpha}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha-\delta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{\delta}(a f, -X), \end{multline*}
where $\epsilon = N_{-\delta,\alpha}$. 
The first 7 factors in the right-hand side are generators of type K1, Z, K1, K2, K1, K1, Z.
The remaining last factor $z_\delta(af, -X)$ can be rewritten as $x_{\delta}(af) \cdot c_{\delta}(-af, X)$ (cf.~\cref{rem:c}) and hence is a product of generators of type Z and K5. \end{proof}

@@

\subsection{The map \texorpdfstring{$S_\alpha(a, -)$}{S(a,-)} and its properties} \label{sec:S-a}
Throughout this section $(A, M)$ is a local pair and, as before, $\alpha, \beta$ are some fixed roots of $\Phi$ forming a sharp angle.

@@

\begin{lemma} \label{R3_1_1} The operators $T_\alpha(aX^{-1})$ satisfy Steinberg relations $\mathrm{R3}_{1,1}^{\angle}$. \end{lemma}
\begin{proof}
Let $a, b \in A$ and $\alpha$, $\beta$ be a pair of roots forming a sharp angle. Set $\epsilon = N_{\alpha,-\beta}$. From~\cref{K-a-b} and the proof of the first part of~\cref{lem:orbit-action} it follows that \[p(g, h, u) \in j_+\left(K(A[X], M[X])\right) \cap P_\alpha(m) \cap P_{\beta}(m')\text{ for some }m,m' \in M.\]
From~\eqref{T_1} we obtain that
\begin{multline} \label{R3_1_1-computation}
 T_\beta(bX^{-1}) \cdot T_\alpha(aX^{-1}) \left(g, [h], u\right) = \\
 = T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ \left[h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right)\right],\ u(1 + am)\right) = \\ 
 = T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ \left[h'\right],\ u(1 + am)\right),
 \end{multline}
where $h' = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am').$
Notice that by~\cref{SR:sharp}
\newline\scalebox{0.85}{\begin{minipage}{1.15\linewidth}\[ p\left(x_\alpha(aX^{-1})\cdot g, h', u(1 + am)\right) = S_\alpha(a, p(g, h, u)) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \in P_\beta(\tfrac{m'}{1+am}), \]\end{minipage}}\vskip\belowdisplayshortskip
therefore we can invoke~\eqref{T_1} once again and can continue~\eqref{R3_1_1-computation} as follows:
\newline\scalebox{0.88}{\begin{minipage}{1.13\linewidth}
\begin{equation}\label{R3_1_1-computation2} \ldots = \left(x_\beta(bX^{-1}) \cdot x_\alpha(aX^{-1}) \cdot g,\ \left[h' \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + \tfrac{bm'}{1+am}}\right)\right],\ u(1 + am + bm')\right). \end{equation}\end{minipage}}\vskip \belowdisplayshortskip
The second component of the above triple can be simplified using~\eqref{eq:conj-h-x} as follows:
\newline\scalebox{0.96}{\begin{minipage}{1.04\linewidth}
\begin{multline*}
 \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha}\left(\tfrac{m'abX^{-1}(1+am)}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am} + \tfrac{m'abX^{-1}}{(1+am)(1 + am + bm')}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right) \cdot h^{-1}_\alpha(1 + am)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right)\right].
\end{multline*}\end{minipage}}\vskip \belowdisplayshortskip
Thus, we see that the expression in the right-hand side of~\eqref{R3_1_1-computation2} would remain unchanged if we swapped $(a,\alpha,m)$ with $(b,\beta,m')$. This implies the required assertion.
\end{proof}

>>>>

The following proposition is the main result of this subsection and also a direct generalization of~\cite[Lemma~3.3]{Tu83}.
\begin{prop} \label{lemma33} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the homomorphism $i_n\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$, induced by the natural embedding of generators, is an isomorphism. Consequently, the obvious homomorphism $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{prop}
\begin{proof}
 We need to construct a homomorphism $j_n$ which would be the inverse of $i_n$. 
 We start with a homomorphism $\widetilde{j}_n^\Phi \colon F\langle \mathcal{X}^\Phi_{\leq n+1} \rangle \to \St^{\leq n}(\Phi, B)$ defined via
 \[ \widetilde{j}^{\Phi}_n(x_\alpha(at^k)) = \begin{cases} x_\alpha(at^k), & k\leq n; \\
      [x_{\alpha - \beta} (N_{\alpha-\beta, \beta} \cdot at^{k-1}), x_{\beta}(t)], & k = n+1, \end{cases} \]
 where $\beta$ is any root of $\Phi$ forming an acute angle with $\alpha$.
 A standard argument, cf. \cite[Proposition~1.1]{Re75} or~\cite[Proposition~3.2.2]{RS76}, shows that $\widetilde{j}^\Phi_n$ does not depend on the choice of $\beta$.

@@

For the rest of this subsection $M$ is the maximal ideal $A$.
We denote by $\beta$ another fixed root of $\Phi$ forming an acute angle with $\alpha$ (i.\,e. $\langle \alpha, \beta \rangle = 1$). We denote by $\Psi$ the subsystem of type $\rA_2$ generated by $\alpha$ and $\beta$.
\begin{df}\label{def:Kab} 
Fix a root $\delta \in \Phi \setminus \Psi$. Define $\widetilde{K}(\alpha, \beta)$ as the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following 5 families of elements parametrized by $\xi \in A[X]$, $f\in M[X]$:
 \begin{enumerate}[label=(K\arabic*)]
  \item $z_\gamma(Xf, \xi)$, for all $\gamma \in \Phi \setminus \Psi$;
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $x_{\alpha-\beta}(Xf)$, $x_{\beta-\alpha}(Xf)$;
  \item $c_{\delta}(f, X\xi)$. \end{enumerate} \end{df}
\begin{prop} \label{K-a-b} One has \[K(A[X], M[X]) = \widetilde{K}(\alpha, \beta) \cdot X_{-\alpha}(M \cdot X) \cdot X_{-\beta}(M \cdot X).\] 
\end{prop} 
\begin{proof}
It is clear that $K(A[X], M[X]) \supseteq \widetilde{K}(\alpha, \beta) \cdot X_{-\alpha}(M \cdot X) \cdot X_{-\beta}(M \cdot X)$. Let us prove the reverse inclusion. Fix an element $g\in K(A[X], M[X])$. We need to show that $g$ lies in the specified product of subgroups.

@@

It remains to verify the assertion for the generators $x_{-\alpha}(X^2f) $ of type K2.
By~\cref{K-a-b-cor} we may assume that the root $\delta$ in~\cref{def:Kab} forms an acute angle with $\alpha$, i.\,e. $\alpha-\delta \in \Phi$. Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = \delta-\alpha$, $\beta = -\delta$ into~\cref{Zrels}\ref{Z5} we obtain that
\begin{multline*} z_{-\alpha}(X^2f, aX^{-1}) = x_{\delta-\alpha}(\epsilon Xf) \cdot x_{\delta}(-af) \cdot x_{-\delta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{\delta-\alpha}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha-\delta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{\delta}(a f, -X), \end{multline*}
where $\epsilon = N_{-\delta,\alpha}$. 
The first 7 factors in the right-hand side are generators of type K1, Z, K1, K2, K1, K1, Z.
The remaining last factor $z_\delta(af, -X)$ can be rewritten as $x_{\delta}(af) \cdot c_{\delta}(-af, X)$ (cf.~\cref{rem:c}) and hence is a product of generators of type Z and K5. \end{proof}

@@

\subsection{The map \texorpdfstring{$S_\alpha(a, -)$}{S(a,-)} and its properties} \label{sec:S-a}
Throughout this section $(A, M)$ is a local pair and, as before, $\alpha, \beta$ are some fixed roots of $\Phi$ forming an acute angle.

@@

\begin{lemma} \label{R3_1_1} The operators $T_\alpha(aX^{-1})$ satisfy Steinberg relations $\mathrm{R3}_{1,1}^{\angle}$. \end{lemma}
\begin{proof}
Let $a, b \in A$ and $\alpha$, $\beta$ be a pair of roots forming an acute angle. Set $\epsilon = N_{\alpha,-\beta}$. From~\cref{K-a-b} and the proof of the first part of~\cref{lem:orbit-action} it follows that \[p(g, h, u) \in j_+\left(K(A[X], M[X])\right) \cap P_\alpha(m) \cap P_{\beta}(m')\text{ for some }m,m' \in M.\]
From~\eqref{T_1} we obtain that
\begin{multline} \label{R3_1_1-computation}
 T_\beta(bX^{-1}) \cdot T_\alpha(aX^{-1}) \left(g, [h], u\right) = \\
 = T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ \left[h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right)\right],\ u(1 + am)\right) = \\ 
 = T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ \left[h'\right],\ u(1 + am)\right),
 \end{multline}
where $h' = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am').$
Notice that by~\cref{SR:sharp}
\newline\scalebox{0.85}{\begin{minipage}{1.15\linewidth}\[ p\left(x_\alpha(aX^{-1})\cdot g, h', u(1 + am)\right) = S_\alpha(a, p(g, h, u)) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \in P_\beta(\tfrac{m'}{1+am}), \]\end{minipage}}\vskip\belowdisplayshortskip
therefore we can invoke~\eqref{T_1} once again and can continue~\eqref{R3_1_1-computation} as follows:
\newline\scalebox{0.88}{\begin{minipage}{1.13\linewidth}
\begin{equation}\label{R3_1_1-computation2} \ldots = \left(x_\beta(bX^{-1}) \cdot x_\alpha(aX^{-1}) \cdot g,\ \left[h' \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + \tfrac{bm'}{1+am}}\right)\right],\ u(1 + am + bm')\right). \end{equation}\end{minipage}}\vskip \belowdisplayshortskip
The second component of the above triple can be simplified using~\eqref{eq:conj-h-x} as follows:
\newline\scalebox{0.96}{\begin{minipage}{1.04\linewidth}
\begin{multline*}
 \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha}\left(\tfrac{m'abX^{-1}(1+am)}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am} + \tfrac{m'abX^{-1}}{(1+am)(1 + am + bm')}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right) \cdot h^{-1}_\alpha(1 + am)\right] = \\
 = \left[ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right)\right].
\end{multline*}\end{minipage}}\vskip \belowdisplayshortskip
Thus, we see that the expression in the right-hand side of~\eqref{R3_1_1-computation2} would remain unchanged if we swapped $(a,\alpha,m)$ with $(b,\beta,m')$. This implies the required assertion.
\end{proof}
